worldmap

### 2024-12-10 analysis/figures/visualization.Rmd
Get rid of the scaling, annotate with new x axes, make it look pretty

slopes <- model_output %>% 
  filter(term == "year_scaled", model_id == "base",
         !metric %in% c("FDiv", "FRic", "FEve", "Jaccard_next")) %>% 
  select(metric, estimate, p.adjust, study_id) %>%
  mutate(estimate_backtrans = case_when(
           metric %in% log_metrics ~ exp(estimate),
           metric %in% logone_metrics ~ 1 + exp(estimate),
           .default = estimate
         )) %>%
  group_by(metric) %>%
  mutate(estimate_scale = scale(estimate_backtrans)[,1]) %>%
  ungroup() %>%
    mutate(ind = case_when(
    estimate > 0 & p.adjust < 0.05 ~ "blue", 
    estimate < 0 & p.adjust < 0.05 ~ "red", 
    .default = "grey"
  )) %>%
  left_join(durations) %>%
  mutate(duration_fac = as.factor(duration)) %>%
  filter(estimate_scale < 3.5, estimate_scale > -3.5)

@@

@@

  metric_plot <- plot_data %>% 
    filter(!is.na(study_id)) %>% 
    ggplot() +
    geom_dotplot(aes(x = estimate_scale, fill = duration_fac),
                 stackgroups = TRUE,
                 method = "histodot",
                 binwidth = 0.1, dotsize = 1.4
    ) +
    scale_fill_manual(values = graid_pal, guide = "none", drop = FALSE) +
    geom_vline(aes(xintercept = gen_est$estimate_scale), linetype = "longdash") +
    #coord_fixed(ratio=5)+#, ylim = c(0,5)) +
    #scale_y_continuous(limits=c(0, 1), expand = c(0, 0), breaks = seq(0, 1,1/10), labels=seq(0,10)) +
    theme_minimal() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          axis.title.y = element_blank(), 
          axis.text.y=element_blank(), 
          axis.title=element_text(size=14)) +
    xlim(-3.5, 3.5) +
    xlab("temporal slope estimate") +
    annotate('text', x = -2.5, y = 0.8, label = plot_label, size = unit(6, "pt"))

  if(isFALSE(keep_x_axis)) {
    metric_plot <- metric_plot + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
  } else{
    metric_plot <- metric_plot + xlim(-3.5, 3.5) +
      geom_hline(aes(yintercept = 0))
  }

  if(gen_est$p.adjust < 0.05){
    metric_plot <- metric_plot + annotate('text', x = (gen_est$estimate_scale + .4), y = 0.8, label = "*", size = unit(16, "pt"))
  }

@@

slope_plot_df <- data.frame(metric_name = c("S", "Jaccard_base", 
                     "SES_FRic", "SES_FEve", "SES_FDiv", 
                     "CWM_diet_inv", "CWM_diet_scav", "CWM_diet_vfish", "CWM_diet_vunk", "CWM_diet_planto",
                     "CWM_diet_seed", "CWM_diet_vend", "CWM_diet_nect", "CWM_diet_fruit", "CWM_diet_vect", 
                     "CWM_forstrat_ground", "CWM_forstrat_watbelowsurf", "CWM_forstrat_wataroundsurf", 
                     "CWM_forstrat_understory", "CWM_forstrat_aerial", 
                     "CWM_forstrat_canopy", "CWM_forstrat_midhigh", "CWM_bodymass_value", rep("empty", 2)),
                     plot_label = c("Species\nRichness", "Jaccard\nSimilarity", 
                     "FRic SES", "FEve SES", "FDiv SES", 
                     "Invertebrates", "Scavenge", "Fish", "Other\nVertebrates", "Other Plants",
                     "Seeds", "Mammals\n& Birds", "Nectar", "Fruit", "Reptiles &\nAmphibians", 
                     "Ground", "Below Water\nSurface", "Around Water\nSurface", 
                     "Understory", "Aerial", "Canopy", "Midhigh", 
                     "Body Mass", rep("empty", 2)), 
                     keep_x_axis = c(FALSE, FALSE, 
                     FALSE, FALSE, TRUE, 
                     FALSE, FALSE, FALSE, FALSE, TRUE,
                     FALSE, FALSE, FALSE, FALSE, TRUE, 
                     FALSE, FALSE, FALSE, 
                     FALSE, TRUE, 
                     FALSE, FALSE, TRUE, rep(FALSE, 2)),
                     color = c(rep(muted[1], 2), rep(muted[2], 3), rep(muted[3], 10), rep(muted[4], 7), rep(muted[5], 3))) %>%
  mutate(across(keep_x_axis, as.logical))

@@

cowplot::save_plot(here::here("figures/slope_plot_grid.png"), slope_plot_grid, ncol = 7, nrow = 5, bg = "white", base_asp = .9)

>>>>

slopes <- model_output %>% 
  filter(term == "year_scaled", model_id == "base",
         !metric %in% c("FDiv", "FRic", "FEve", "Jaccard_next", "Jaccard_hind")) %>% 
  select(metric, estimate, p.adjust, study_id) %>%
  mutate(estimate_backtrans = case_when(
           metric %in% log_metrics ~ exp(estimate),
           metric %in% logone_metrics ~ 1 + exp(estimate),
           .default = estimate
         )) %>%
  group_by(metric) %>%
  mutate(estimate_scale = scale(estimate_backtrans)[,1]) %>%
  ungroup() %>%
    mutate(ind = case_when(
    estimate > 0 & p.adjust < 0.05 ~ "blue", 
    estimate < 0 & p.adjust < 0.05 ~ "red", 
    .default = "grey"
  )) %>%
  left_join(durations) %>%
  mutate(duration_fac = as.factor(duration)) %>%
  filter(estimate_scale < 3.5, estimate_scale > -3.5)

@@

  # plotting parameters
  lim_buffer <- plot_data %>% filter(!is.na(estimate)) %>% pull(estimate) %>% sd()/4
  lim <- max(abs(plot_data$estimate))

@@

  metric_plot <- plot_data %>% 
    filter(!is.na(study_id)) %>% 
    ggplot() +
    geom_dotplot(aes(x = estimate, fill = duration_fac),
                 stackgroups = TRUE,
                 method = "histodot"#,
                 #binwidth = 0.1, dotsize = 1.4
    ) +
    scale_fill_manual(values = graid_pal, guide = "none", drop = FALSE) +
    geom_vline(aes(xintercept = gen_est$estimate), linetype = "longdash") +
    #coord_fixed(ratio=5)+#, ylim = c(0,5)) +
    #scale_y_continuous(limits=c(0, 1), expand = c(0, 0), breaks = seq(0, 1,1/10), labels=seq(0,10)) +
    theme_minimal() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          axis.title.y = element_blank(), 
          axis.text.y=element_blank()#, 
          #axis.title=element_text(size=14)
          ) +
    xlim(-lim - lim_buffer, 
         lim + lim_buffer) +
    xlab("temporal slope estimate") +
    geom_hline(aes(yintercept = 0)) +
    annotate('text', x = (-lim + (2*lim_buffer)), y = 0.8, label = plot_label, size = unit(4, "pt"))

  if(isFALSE(keep_x_axis)) {
    #metric_plot <- metric_plot + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
  } else{
    #metric_plot <- metric_plot + xlim(-3.5, 3.5) +
      #geom_hline(aes(yintercept = 0))
  }

  if(gen_est$p.adjust < 0.05){
    metric_plot <- metric_plot + annotate('text', x = (gen_est$estimate + (lim_buffer*2)), y = 0.8, label = "*", size = unit(16, "pt"))
  }

@@

slope_plot_df <- data.frame(metric_name = c("S", "Jaccard_base", 
                     "SES_FRic", "SES_FEve", "SES_FDiv", 
                     "CWM_diet_inv", "CWM_diet_scav", "CWM_diet_vfish", "CWM_diet_vunk", "CWM_diet_planto",
                     "CWM_diet_seed", "CWM_diet_vend", "CWM_diet_nect", "CWM_diet_fruit", "CWM_diet_vect", 
                     "CWM_forstrat_ground", "CWM_forstrat_watbelowsurf", "CWM_forstrat_wataroundsurf", 
                     "CWM_forstrat_understory", "CWM_forstrat_aerial", 
                     "CWM_forstrat_canopy", "CWM_forstrat_midhigh", "CWM_bodymass_value", rep("empty", 2)),
                     plot_label = c("Species\nRichness", "Jaccard\nSimilarity", 
                     "FRic SES", "FEve SES", "FDiv SES", 
                     "Invertebrates", "Scavenge", "Fish", "Other\nVertebrates", "Other\nPlants",
                     "Seeds", "Mammals\n& Birds", "Nectar", "Fruit", "Reptiles &\nAmphibians", 
                     "Ground", "Below Water\nSurface", "Around\nWater\nSurface", 
                     "Understory", "Aerial", "Canopy", "Midhigh", 
                     "Body Mass", rep("empty", 2)), 
                     keep_x_axis = c(FALSE, FALSE, 
                     FALSE, FALSE, TRUE, 
                     FALSE, FALSE, FALSE, FALSE, TRUE,
                     FALSE, FALSE, FALSE, FALSE, TRUE, 
                     FALSE, FALSE, FALSE, 
                     FALSE, TRUE, 
                     FALSE, FALSE, TRUE, rep(FALSE, 2)),
                     color = c(rep(muted[1], 2), rep(muted[2], 3), rep(muted[3], 10), rep(muted[4], 7), rep(muted[5], 3))) %>%
  mutate(across(keep_x_axis, as.logical))

@@

cowplot::save_plot(here::here("figures/slope_plot_grid_test.png"), slope_plot_grid, ncol = 7, nrow = 5, bg = "white", base_asp = .7)
